---
title: 'Jonah Crab CCM significance testing'
date-modified: '2023-07-28'
execute: 
  message: false
  warning: false
editor_options: 
  chunk_output_type: console
editor: 
  markdown: 
    wrap: 72
---

# Re-run CCM with max lib size

## Aggregate (done)

### Catch
```{r}
# Aggregate - CATCH max lib size
RESULTS_ccm_agg_max <- pmap_dfr(params_ccm_combos,function(jonah,prey){
  df_temp <- catchCCMdf_agg
  lib_vec <- paste(1, nrow(df_temp))
  rho_E_1<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = jonah,target = prey, maxE = 7, showPlot = FALSE)
  E_out_1<-rho_E_1[which.max(rho_E_1$rho),"E"][1]
  out_1 <- CCM(dataFrame= df_temp, columns=jonah, target=prey, E = E_out_1, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_1, nrow(df_temp) - E_out_1, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE) %>%
    mutate(jonah=jonah,prey=prey, direction= paste("jonah","->","prey"),E = E_out_1)
  rho_E_2<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = prey,target = jonah, maxE = 7, showPlot = FALSE)
  E_out_2<-rho_E_2[which.max(rho_E_1$rho),"E"][1]
  out_2 <- CCM(dataFrame= df_temp, columns=prey, target=jonah, E = E_out_2, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_2, nrow(df_temp)-E_out_2, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE)  %>%
   mutate(jonah=jonah,prey=prey,direction= paste("prey","->","jonah"),E = E_out_2)
  bind_rows(out_1,out_2)
}) %>% addDirection()

RESULTS_ccm_agg_max<- RESULTS_ccm_agg_max %>% group_by(xmap) %>% pivot_longer(cols=all_of(combos)) %>% na.omit() %>% filter(substr(xmap, 1, 1)==substr(name, 1, 1))

```

### Weight
```{r}
# Aggregate - CATCH max lib size
RESULTS_ccm_wt_agg_max <- pmap_dfr(params_ccm_combos,function(jonah,prey){
  df_temp <- wtCCMdf_agg
  lib_vec <- paste(1, nrow(df_temp))
  rho_E_1<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = jonah,target = prey, maxE = 7, showPlot = FALSE)
  E_out_1<-rho_E_1[which.max(rho_E_1$rho),"E"][1]
  out_1 <- CCM(dataFrame= df_temp, columns=jonah, target=prey, E = E_out_1, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_1, nrow(df_temp) - E_out_1, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE) %>%
    mutate(jonah=jonah,prey=prey, direction= paste("jonah","->","prey"),E = E_out_1)
  rho_E_2<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = prey,target = jonah, maxE = 7, showPlot = FALSE)
  E_out_2<-rho_E_2[which.max(rho_E_1$rho),"E"][1]
  out_2 <- CCM(dataFrame= df_temp, columns=prey, target=jonah, E = E_out_2, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_2, nrow(df_temp)-E_out_2, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE)  %>%
   mutate(jonah=jonah,prey=prey,direction= paste("prey","->","jonah"),E = E_out_2)
  bind_rows(out_1,out_2)
}) %>% addDirection()

RESULTS_ccm_wt_agg_max<- RESULTS_ccm_wt_agg_max %>% group_by(xmap) %>% pivot_longer(cols=all_of(combos)) %>% na.omit() %>% filter(substr(xmap, 1, 1)==substr(name, 1, 1))

```


## Regions (done)

### Catch
```{r}
# By region, each individually - CATCH max lib size
RESULTS_ccm_by_reg_max <- pmap_dfr(params_regions_combos,function(jonah,prey, region){
  df_temp <- catchCCMdf_reg %>% filter(Region == region)
  lib_vec <- paste(1, nrow(df_temp))
  rho_E_1<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = jonah,target = prey, maxE = 7, showPlot = FALSE)
  E_out_1<-rho_E_1[which.max(rho_E_1$rho),"E"][1]
  out_1 <- CCM(dataFrame= df_temp, columns=jonah, target=prey, E = E_out_1, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_1, nrow(df_temp) - E_out_1, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE) %>%
    mutate(jonah=jonah,prey=prey, direction= paste("jonah","->","prey"),region=region,E = E_out_1)
  rho_E_2<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = prey,target = jonah, maxE = 7, showPlot = FALSE)
  E_out_2<-rho_E_2[which.max(rho_E_1$rho),"E"][1]
  out_2 <- CCM(dataFrame= df_temp, columns=prey, target=jonah, E = E_out_2, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_2, nrow(df_temp)-E_out_2, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE)  %>%
   mutate(jonah=jonah,prey=prey,direction= paste("prey","->","jonah"),region=region,E = E_out_2)
  bind_rows(out_1,out_2)
}) %>% addDirection()

RESULTS_ccm_by_reg_max<- RESULTS_ccm_by_reg_max %>% group_by(xmap) %>% pivot_longer(cols=all_of(combos)) %>% na.omit() %>% filter(substr(xmap, 1, 1)==substr(name, 1, 1))

```

### Weight
```{r}
# By region, each individually - WEIGHT max lib size
RESULTS_ccm_wt_by_reg_max <- pmap_dfr(params_regions_combos,function(jonah,prey, region){
  df_temp <- wtCCMdf_reg %>% filter(Region == region)
  lib_vec <- paste(1, nrow(df_temp))
    
  rho_E_1<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = jonah,target = prey, maxE = 7, showPlot = FALSE)
  E_out_1<-rho_E_1[which.max(rho_E_1$rho),"E"][1]
  out_1 <- CCM(dataFrame= df_temp, columns=jonah, target=prey, E = E_out_1, Tp=1,
               libSizes = paste(nrow(df_temp)-E_out_1, nrow(df_temp) - E_out_1, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE) %>%
    mutate(jonah=jonah,prey=prey, direction= paste("jonah","->","prey"),region=region,E = E_out_1)
  rho_E_2<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = prey,target = jonah, maxE = 7, showPlot = FALSE)
  E_out_2<-rho_E_2[which.max(rho_E_1$rho),"E"][1]
  out_2 <- CCM(dataFrame= df_temp, columns=prey, target=jonah, E = E_out_2, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_2, nrow(df_temp)-E_out_2, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE)  %>%
    mutate(jonah=jonah,prey=prey,direction= paste("prey","->","jonah"),region=region,E = E_out_2)
  bind_rows(out_1,out_2)
}) %>% addDirection()

RESULTS_ccm_wt_by_reg_max<- RESULTS_ccm_wt_by_reg_max %>% group_by(xmap) %>% pivot_longer(cols=all_of(combos)) %>% na.omit() %>% filter(substr(xmap, 1, 1)==substr(name, 1, 1))

```

## Strata (done)
### Catch
```{r}
RESULTS_ccm_by_strat_max <- pmap_dfr(params_strat_combos,function(jonah,prey, stratum){
  df_temp <- catchCCMdf_strat %>% filter(Stratum == stratum)
  lib_vec <- paste(1, nrow(df_temp))
  rho_E_1<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = jonah,target = prey, maxE = 7, showPlot = FALSE)
  E_out_1<-rho_E_1[which.max(rho_E_1$rho),"E"][1]
  out_1 <- CCM(dataFrame= df_temp, columns=jonah, target=prey, E = E_out_1, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_1, nrow(df_temp) - E_out_1, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE) %>%
    mutate(jonah=jonah,prey=prey, direction= paste("jonah","->","prey"),stratum=stratum,E = E_out_1)
  rho_E_2<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = prey,target = jonah, maxE = 7, showPlot = FALSE)
  E_out_2<-rho_E_2[which.max(rho_E_1$rho),"E"][1]
  out_2 <- CCM(dataFrame= df_temp, columns=prey, target=jonah, E = E_out_2, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_2, nrow(df_temp)-E_out_2, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE)  %>%
   mutate(jonah=jonah,prey=prey,direction= paste("prey","->","jonah"),stratum=stratum,E = E_out_2)
  bind_rows(out_1,out_2)
}) %>% addDirection()

RESULTS_ccm_by_strat_max<- RESULTS_ccm_by_strat_max %>% group_by(xmap) %>% pivot_longer(cols=all_of(combos)) %>% na.omit() %>% filter(substr(xmap, 1, 1)==substr(name, 1, 1))

```

### Weight
```{r}
RESULTS_ccm_wt_by_strat_max <- pmap_dfr(params_strat_combos,function(jonah,prey, stratum){
  df_temp <- wtCCMdf_strat %>% filter(Stratum == stratum)
  lib_vec <- paste(1, nrow(df_temp))
  rho_E_1<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = jonah,target = prey, maxE = 7, showPlot = FALSE)
  E_out_1<-rho_E_1[which.max(rho_E_1$rho),"E"][1]
  out_1 <- CCM(dataFrame= df_temp, columns=jonah, target=prey, E = E_out_1, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_1, nrow(df_temp) - E_out_1, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE) %>%
    mutate(jonah=jonah,prey=prey, direction= paste("jonah","->","prey"),stratum=stratum,E = E_out_1)
  rho_E_2<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = prey,target = jonah, maxE = 7, showPlot = FALSE)
  E_out_2<-rho_E_2[which.max(rho_E_1$rho),"E"][1]
  out_2 <- CCM(dataFrame= df_temp, columns=prey, target=jonah, E = E_out_2, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_2, nrow(df_temp)-E_out_2, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE)  %>%
   mutate(jonah=jonah,prey=prey,direction= paste("prey","->","jonah"),stratum=stratum,E = E_out_2)
  bind_rows(out_1,out_2)
}) %>% addDirection()

RESULTS_ccm_wt_by_strat_max<- RESULTS_ccm_wt_by_strat_max %>% group_by(xmap) %>% pivot_longer(cols=all_of(combos)) %>% na.omit() %>% filter(substr(xmap, 1, 1)==substr(name, 1, 1))
```


## Areas (done)

### Catch

```{r}
RESULTS_ccm_by_area_max <- pmap_dfr(params_areas_ccm_combos,function(jonah,prey, areaInput){
  df_temp <- catchCCMdf %>% filter(area==areaInput)
  lib_vec <- paste(1, nrow(df_temp))
  rho_E_1<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = jonah,target = prey, maxE = 7, showPlot = FALSE)
  E_out_1<-rho_E_1[which.max(rho_E_1$rho),"E"][1]
  out_1 <- CCM(dataFrame= df_temp, columns=jonah, target=prey, E = E_out_1, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_1, nrow(df_temp) - E_out_1, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE) %>%
    mutate(jonah=jonah,prey=prey, direction= paste("jonah","->","prey"),areaInput=areaInput,E = E_out_1)
  rho_E_2<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = prey,target = jonah, maxE = 7, showPlot = FALSE)
  E_out_2<-rho_E_2[which.max(rho_E_1$rho),"E"][1]
  out_2 <- CCM(dataFrame= df_temp, columns=prey, target=jonah, E = E_out_2, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_2, nrow(df_temp)-E_out_2, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE)  %>%
   mutate(jonah=jonah,prey=prey,direction= paste("prey","->","jonah"),areaInput=areaInput,E = E_out_2)
  bind_rows(out_1,out_2)
}) %>% addDirection()

RESULTS_ccm_by_area_max<- RESULTS_ccm_by_area_max %>% group_by(xmap) %>% pivot_longer(cols=all_of(combos)) %>% na.omit() %>% filter(substr(xmap, 1, 1)==substr(name, 1, 1))
```


### Weight
```{r}
RESULTS_ccm_wt_by_area_max <- pmap_dfr(params_areas_ccm_combos,function(jonah,prey, areaInput){
  df_temp <- wtCCMdf %>% filter(area==areaInput)
  lib_vec <- paste(1, nrow(df_temp))
  rho_E_1<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = jonah,target = prey, maxE = 7, showPlot = FALSE)
  E_out_1<-rho_E_1[which.max(rho_E_1$rho),"E"][1]
  out_1 <- CCM(dataFrame= df_temp, columns=jonah, target=prey, E = E_out_1, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_1, nrow(df_temp) - E_out_1, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE) %>%
    mutate(jonah=jonah,prey=prey, direction= paste("jonah","->","prey"),areaInput=areaInput,E = E_out_1)
  rho_E_2<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = prey,target = jonah, maxE = 7, showPlot = FALSE)
  E_out_2<-rho_E_2[which.max(rho_E_1$rho),"E"][1]
  out_2 <- CCM(dataFrame= df_temp, columns=prey, target=jonah, E = E_out_2, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_2, nrow(df_temp)-E_out_2, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE)  %>%
   mutate(jonah=jonah,prey=prey,direction= paste("prey","->","jonah"),areaInput=areaInput,E = E_out_2)
  bind_rows(out_1,out_2)
}) %>% addDirection()

RESULTS_ccm_wt_by_area_max<- RESULTS_ccm_wt_by_area_max %>% group_by(xmap) %>% pivot_longer(cols=all_of(combos)) %>% na.omit() %>% filter(substr(xmap, 1, 1)==substr(name, 1, 1))
```



The second criteria for significant convergence is statistically significant cross-map skill at maximum library size. Although we got p-values for max lib size above using the compute_stats function, the validity of these p-values is questionable. The standard method of determining significance is with randomization tests, where you randomly rearrange the predictor time series to destroy the dynamic relationship between variables and perform CCM between the shuffled time series and the original target time series. This randomization is repeated 100 (or 1000) times to generate a null distribution of cross-map skill, against which the actual value can be compared.

We will use the "surrogate" function from the package tseries. We'll start with the aggregate catch CCM.

# Surrogate data

## Aggregate (done)
### Catch
```{r}
#Create 200 random surrogates of the jonah column and bind to the date, scallop, and rock columns of the data frame we used for aggregate CCM
set.seed(7)
ccm_agg_surr <- cbind(catchCCMdf_agg, surrogate(catchCCMdf_agg$jonah, ns=200)) %>% select(-jonah)

#Rename the surrogate columns to be of the form "jonah_1", "jonah_2", etc.
ccm_agg_surr <- rename_with(ccm_agg_surr,.cols=!any_of(c("date", "scallop", "rock")),~ paste0("jonah_", .x, recycle0 = TRUE))

#Create the data frame that will store the combinations we want to test
params_ccm_combos_surr <- map_dfr(seq_len(200), ~params_ccm_combos %>% select(-"jonah"))
params_ccm_combos_surr <- params_ccm_combos_surr %>% mutate(surr_trial = rep(1:200, each=2))
```

```{r}

RESULTS_ccm_agg_surr <- pmap_dfr(params_ccm_combos_surr,function(prey, surr_trial){
  lib_vec <- paste(1, nrow(ccm_agg_surr))
  jonah_col = paste("jonah",surr_trial, sep="_" )
  
  #The only difference for finding E opt is that the column name for jonah is now jonah_n, 
  #where n is the surrogate trial number
  rho_E_1<- EmbedDimension(dataFrame = ccm_agg_surr, lib = lib_vec, pred = lib_vec, 
                           columns = jonah_col,target = prey, maxE = 7, showPlot = FALSE) 
  E_out_1<-rho_E_1[which.max(rho_E_1$rho),"E"][1] #store E
  
  #Run CCM - jonah to rock/scallop
  #We are only running the CCM at max lib size
  out_1 <- CCM(dataFrame= ccm_agg_surr, columns=jonah_col, target=prey, E = E_out_1, Tp=1,
               libSizes = paste(nrow(ccm_agg_surr) - E_out_1, nrow(ccm_agg_surr) - E_out_1, 
                                "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE) %>%
    mutate(prey=prey,direction= paste("jonah","->","prey"),
           E = E_out_1, trial_num = surr_trial) %>% 
  rename_with( ~ paste0("jonah:", prey, recycle0 = TRUE), starts_with("jonah")) %>% 
  rename_with( ~ paste0(prey, ":jonah",recycle0 = TRUE), starts_with(prey))
  
  #find optimal E for predicting jonah from scallop or rock (i.e., scallop/rock -> jonah)
  rho_E_2<- EmbedDimension(dataFrame = ccm_agg_surr, lib = lib_vec, pred = lib_vec, 
                           columns = prey,target = jonah_col, maxE = 7, showPlot = FALSE)
  E_out_2<-rho_E_2[which.max(rho_E_1$rho),"E"][1] #store E
  
  #Run CCM - rock/scallop to jonah
  out_2 <- CCM(dataFrame= ccm_agg_surr, columns=prey, target=jonah_col, E = E_out_2, Tp=1, 
               libSizes = paste(nrow(ccm_agg_surr)-E_out_2, nrow(ccm_agg_surr)-E_out_2, "1",sep=" "), 
               sample=1, verbose=FALSE, showPlot = FALSE) %>%
    mutate(prey=prey, direction= paste("prey","->","jonah"),
           E = E_out_2, trial_num = surr_trial)  %>% 
  rename_with( ~ paste0("jonah:", prey, recycle0 = TRUE), starts_with("jonah")) %>% 
  rename_with( ~ paste0(prey, ":jonah",recycle0 = TRUE), starts_with(prey))
  
  bind_rows(out_1,out_2)
  
}) %>% mutate("jonah"=NA) %>% addDirection()
```

Did we get a null distribution? Yes!
```{r}
js_c_null <- RESULTS_ccm_agg_surr %>% group_by(xmap) %>% 
  pivot_longer(cols=all_of(combos)) %>% 
  na.omit() %>% 
  filter(substr(xmap, 1, 1)==substr(name, 1, 1))  %>% 
  filter(name=="jonah:scallop")

quantile(js_c_null$value, .95)
1-ecdf(js_c_null$value)(0.237) #p=0.26

ggplot(js_c_null, aes(value)) +
stat_ecdf(geom = "step")+
  geom_hline(yintercept = 0.95)+
 # geom_vline(xintercept = quantile(null_dist(df = RESULTS_ccm_wt_by_reg_surr %>% filter(region==1), s="scallop:jonah"), .95))+
  geom_vline(xintercept = 0.237, color="red")
```

Unfortunately, we can see that our empirical cross-map skill of 0.237 does not exceed 0.432, the p-value threshold of 0.05; 26% of the surrogate trials produced a cross-map skill equal to or higher than 0.225.
Let's write a function to quickly check the others:

```{r}
null_dist <- function(df, s) {
  df_out <- df %>% group_by(xmap) %>% 
  pivot_longer(cols=all_of(combos)) %>% 
  na.omit() %>% 
  filter(substr(xmap, 1, 1)==substr(name, 1, 1)) %>% 
  filter(name==s)
  return(df_out %>% pull(value))
}

null_dist(RESULTS_ccm_agg_surr, s="jonah:scallop")
1-ecdf(null_dist(RESULTS_ccm_agg_surr, s="jonah:scallop"))(0.237) #p=0.26

```

```{r}
find_p_agg <- function(y) {
  ndist = null_dist(df = RESULTS_ccm_agg_surr, s=y)
    val = RESULTS_ccm_agg_max %>% filter(name==y) %>% pull(value)
    prob = (sum(ndist >= val))/200
    return(prob)
}

data.frame(combos, stringsAsFactors = FALSE) %>% rowwise() %>% mutate(p = find_p_agg(combos))

```

### Weight
```{r}
#Create 200 random surrogates of the jonah column and bind to the date, scallop, and rock columns of the data frame we used for aggregate CCM
set.seed(7)
ccm_wt_agg_surr <- cbind(wtCCMdf_agg, surrogate(wtCCMdf_agg$jonah, ns=200)) %>% select(-jonah)

#Rename the surrogate columns to be of the form "jonah_1", "jonah_2", etc.
ccm_wt_agg_surr <- rename_with(ccm_wt_agg_surr,.cols=!any_of(c("date", "scallop", "rock")),~ paste0("jonah_", .x, recycle0 = TRUE))
```

```{r}
RESULTS_ccm_wt_agg_surr <- pmap_dfr(params_ccm_combos_surr,function(prey, surr_trial){
  lib_vec <- paste(1, nrow(ccm_wt_agg_surr))
  jonah_col = paste("jonah",surr_trial, sep="_" )
  
  rho_E_1<- EmbedDimension(dataFrame = ccm_wt_agg_surr, lib = lib_vec, pred = lib_vec, 
                           columns = jonah_col,target = prey, maxE = 7, showPlot = FALSE) 
  E_out_1<-rho_E_1[which.max(rho_E_1$rho),"E"][1] #store E

  out_1 <- CCM(dataFrame= ccm_wt_agg_surr, columns=jonah_col, target=prey, E = E_out_1, Tp=1,
               libSizes = paste(nrow(ccm_wt_agg_surr) - E_out_1, nrow(ccm_wt_agg_surr) - E_out_1, 
                                "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE) %>%
    mutate(prey=prey,direction= paste("jonah","->","prey"),
           E = E_out_1, trial_num = surr_trial) %>% 
  rename_with( ~ paste0("jonah:", prey, recycle0 = TRUE), starts_with("jonah")) %>% 
  rename_with( ~ paste0(prey, ":jonah",recycle0 = TRUE), starts_with(prey))
  
  rho_E_2<- EmbedDimension(dataFrame = ccm_wt_agg_surr, lib = lib_vec, pred = lib_vec, 
                           columns = prey,target = jonah_col, maxE = 7, showPlot = FALSE)
  E_out_2<-rho_E_2[which.max(rho_E_1$rho),"E"][1] #store E
  
  out_2 <- CCM(dataFrame= ccm_wt_agg_surr, columns=prey, target=jonah_col, E = E_out_2, Tp=1, 
               libSizes = paste(nrow(ccm_wt_agg_surr)-E_out_2, nrow(ccm_wt_agg_surr)-E_out_2, "1",sep=" "), 
               sample=1, verbose=FALSE, showPlot = FALSE) %>%
    mutate(prey=prey, direction= paste("prey","->","jonah"),
           E = E_out_2, trial_num = surr_trial)  %>% 
  rename_with( ~ paste0("jonah:", prey, recycle0 = TRUE), starts_with("jonah")) %>% 
  rename_with( ~ paste0(prey, ":jonah",recycle0 = TRUE), starts_with(prey))
  
  bind_rows(out_1,out_2)
  
}) %>% mutate("jonah"=NA) %>% addDirection()
```

```{r}
find_p_agg_wt <- function(y) {
  ndist = null_dist(df = RESULTS_ccm_wt_agg_surr, s=y)
    val = RESULTS_ccm_wt_agg_max %>% filter(name==y) %>% pull(value)
    prob = (sum(ndist >= val))/200
    return(prob)
}

data.frame(combos, stringsAsFactors = FALSE) %>% rowwise() %>% mutate(p = find_p_agg_wt(combos))

```

## By region (done)

### Catch
Create surrogate data
```{r}
set.seed(7)
ccm_reg_surr <- catchCCMdf_reg %>% group_by(Region) %>% mutate(new=surrogate(jonah, ns=200)) %>% select(-jonah) %>% mutate(new=data.frame(new)) %>% unnest(cols=new)

#Rename the surrogate columns to be of the form "jonah_1", "jonah_2", etc.
ccm_reg_surr <- rename_with(ccm_reg_surr,.cols=!any_of(c("date", "scallop", "rock", "Region")),~ paste0("jonah_", .x, recycle0 = TRUE) %>% str_remove("X"))

#Create the data frame that will store the combinations we want to test
params_regions_combos_surr <- map_dfr(seq_len(200), ~params_regions_combos %>% select(-"jonah"))
params_regions_combos_surr <- params_regions_combos_surr %>% mutate(surr_trial = rep(1:200, each=10))
```

```{r}
# Random CCM by region, each individually - CATCH

RESULTS_ccm_by_reg_surr <- pmap_dfr(params_regions_combos_surr,function(prey, region, surr_trial){
  df_temp <- ccm_reg_surr %>% filter(Region == region)
  lib_vec <- paste(1, nrow(df_temp))
  jonah_col = paste("jonah",surr_trial, sep="_" )
  
  rho_E_1<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = jonah_col,target = prey, maxE = 7, showPlot = FALSE)
  E_out_1<-rho_E_1[which.max(rho_E_1$rho),"E"][1]
  out_1 <- CCM(dataFrame= df_temp, columns=jonah_col, target=prey, E = E_out_1, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_1, nrow(df_temp) - E_out_1, 
                                "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE) %>%
    mutate(prey=prey, direction= paste("jonah","->","prey"),
           E = E_out_1,region=region, trial_num = surr_trial) %>% 
    rename_with( ~ paste0("jonah:", prey, recycle0 = TRUE), starts_with("jonah")) %>% 
    rename_with( ~ paste0(prey, ":jonah",recycle0 = TRUE), starts_with(prey))
  
  rho_E_2<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = prey,target = jonah_col, maxE = 7, showPlot = FALSE)
  E_out_2<-rho_E_2[which.max(rho_E_1$rho),"E"][1]
  out_2 <- CCM(dataFrame= df_temp, columns=prey, target=jonah_col, E = E_out_2, Tp=1,
               libSizes = paste(nrow(df_temp)-E_out_2, nrow(df_temp)-E_out_2, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE)  %>%
    mutate(prey=prey,direction= paste("prey","->","jonah"),
           E = E_out_2,region=region, trial_num = surr_trial) %>% 
    rename_with( ~ paste0("jonah:", prey, recycle0 = TRUE), starts_with("jonah")) %>% 
    rename_with( ~ paste0(prey, ":jonah",recycle0 = TRUE), starts_with(prey))
  

  bind_rows(out_1,out_2)
}) %>% mutate("jonah"=NA)  %>% addDirection()
```


```{r}
find_p_reg <- function(x,y) {
  ndist = null_dist(df = RESULTS_ccm_by_reg_surr %>% filter(region==x), s=y)
    val = RESULTS_ccm_by_reg_max %>% filter(region==x, name==y) %>% pull(value)
    prob = (sum(ndist >= val))/200
    return(prob)
}

reg_df <- data.frame(expand.grid(c(1:5),combos, stringsAsFactors = FALSE))
#map2(reg_df$Var1, reg_df$Var2, find_p_reg )
reg_df %>% rowwise() %>% mutate(p = find_p_reg(Var1, Var2))

```

### Weight
```{r}
#Create surrogate data
set.seed(7)
ccm_reg_wt_surr <- wtCCMdf_reg %>% group_by(Region) %>% mutate(new=surrogate(jonah, ns=200)) %>% select(-jonah) %>% mutate(new=data.frame(new)) %>% unnest(cols=new)

#Rename the surrogate columns to be of the form "jonah_1", "jonah_2", etc.
ccm_reg_wt_surr <- rename_with(ccm_reg_wt_surr,.cols=!any_of(c("date", "scallop", "rock", "Region")),~ paste0("jonah_", .x, recycle0 = TRUE) %>% str_remove("X"))

# Random CCM by region, each individually - weight

RESULTS_ccm_wt_by_reg_surr <- pmap_dfr(params_regions_combos_surr,function(prey, region, surr_trial){
  df_temp <- ccm_reg_wt_surr %>% filter(Region == region)
  lib_vec <- paste(1, nrow(df_temp))
  jonah_col = paste("jonah",surr_trial, sep="_" )
  
  rho_E_1<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = jonah_col,target = prey, maxE = 7, showPlot = FALSE)
  E_out_1<-rho_E_1[which.max(rho_E_1$rho),"E"][1]
  out_1 <- CCM(dataFrame= df_temp, columns=jonah_col, target=prey, E = E_out_1, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_1, nrow(df_temp) - E_out_1, 
                                "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE) %>%
    mutate(prey=prey,
           direction= paste("jonah","->","prey"),
           E = E_out_1,region=region, trial_num = surr_trial) %>% 
    rename_with( ~ paste0("jonah:", prey, recycle0 = TRUE), starts_with("jonah")) %>% 
    rename_with( ~ paste0(prey, ":jonah",recycle0 = TRUE), starts_with(prey))
  
  rho_E_2<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = prey,target = jonah_col, maxE = 7, showPlot = FALSE)
  E_out_2<-rho_E_2[which.max(rho_E_1$rho),"E"][1]
  out_2 <- CCM(dataFrame= df_temp, columns=prey, target=jonah_col, E = E_out_2, Tp=1,
               libSizes = paste(nrow(df_temp)-E_out_2, nrow(df_temp)-E_out_2, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE)  %>%
    mutate(prey=prey,
           direction= paste("prey","->","jonah"),
           E = E_out_2,region=region, trial_num = surr_trial) %>% 
    rename_with( ~ paste0("jonah:", prey, recycle0 = TRUE), starts_with("jonah")) %>% 
    rename_with( ~ paste0(prey, ":jonah",recycle0 = TRUE), starts_with(prey))
  bind_rows(out_1,out_2)
}) %>% mutate("jonah"=NA)  %>% addDirection()

```

```{r}
find_p_reg_wt <- function(x,y) {
  ndist = null_dist(df = RESULTS_ccm_wt_by_reg_surr %>% filter(region==x), s=y)
    val = RESULTS_ccm_wt_by_reg_max %>% filter(region==x, name==y) %>% pull(value)
    prob = (sum(ndist >= val))/200
    return(prob)
}

reg_df <- data.frame(expand.grid(c(1:5),combos, stringsAsFactors = FALSE))
#map2(reg_df$Var1, reg_df$Var2, find_p_reg_wt )

reg_df %>% rowwise() %>% mutate(p = find_p_reg_wt(Var1, Var2))

 ggplot(data.frame(x=null_dist(df = RESULTS_ccm_wt_by_reg_surr %>% filter(region==1), s="scallop:jonah")), aes(x)) +
   stat_ecdf(geom = "step")+
   #geom_hline(yintercept = 0.95)+
   geom_vline(xintercept = 0.1289)+
   geom_vline(xintercept = 0.1955, color="red")

# ggplot(data.frame(x=null_dist(df = RESULTS_ccm_wt_by_reg_surr %>% filter(region==2), s="scallop:jonah")), aes(x)) +
#   stat_ecdf(geom = "step")+
#   geom_hline(yintercept = 0.95)+
#   geom_vline(xintercept = 0.1612493)+
#   geom_vline(xintercept = 0.312, color="red")+theme_bw()
# 
 ggplot(data.frame(x=null_dist(df = RESULTS_ccm_wt_by_reg_surr %>% filter(region==1), s="scallop:jonah")), aes(x)) +
   geom_histogram(bins = 40)+
geom_vline(xintercept = quantile(null_dist(df = RESULTS_ccm_wt_by_reg_surr %>% filter(region==1), s="scallop:jonah"), .95))+
   geom_vline(xintercept = 0.1955, color="red")+theme_bw()+labs(x="Cross-map skill", y="Count")+theme(axis.text=element_text(size=10),
        axis.title.x=element_text(size=12, margin = margin(t=10)), axis.title.y = element_text(margin = margin(r=10)))
```

scallop:jonah for weight is highly statistically significant in regions 1 and 2!

Now let's look at strata.
## By stratum (done)

### Catch
```{r}
#Create the data frame that will store the combinations we want to test
params_strat_combos_surr <- map_dfr(seq_len(200), ~params_strat_combos %>% select(-"jonah"))
params_strat_combos_surr <- params_strat_combos_surr %>% mutate(surr_trial = rep(1:200, each=8))

#Create surrogate data
set.seed(7)
ccm_strat_surr <- catchCCMdf_strat %>% group_by(Stratum) %>% mutate(new=surrogate(jonah, ns=200)) %>% select(-jonah) %>% mutate(new=data.frame(new)) %>% unnest(cols=new)

ccm_strat_surr <- rename_with(ccm_strat_surr,.cols=!any_of(c("date", "scallop", "rock", "Stratum")),~ paste0("jonah_", .x, recycle0 = TRUE) %>% str_remove("X"))

```

```{r}
RESULTS_ccm_by_strat_surr <- pmap_dfr(params_strat_combos_surr,function(prey, stratum, surr_trial){
  df_temp <- ccm_strat_surr %>% filter(Stratum==stratum)
  lib_vec <- paste(1, nrow(df_temp))
  jonah_col = paste("jonah",surr_trial, sep="_" )
  
  rho_E_1<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = jonah_col,target = prey, maxE = 7, showPlot = FALSE)
  E_out_1<-rho_E_1[which.max(rho_E_1$rho),"E"][1]
  out_1 <- CCM(dataFrame= df_temp, columns=jonah_col, target=prey, E = E_out_1, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_1, nrow(df_temp) - E_out_1, 
                                "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE) %>%
    mutate(prey=prey, direction= paste("jonah","->","prey"),
           E = E_out_1,stratum=stratum, trial_num = surr_trial) %>% 
    rename_with( ~ paste0("jonah:", prey, recycle0 = TRUE), starts_with("jonah")) %>% 
    rename_with( ~ paste0(prey, ":jonah",recycle0 = TRUE), starts_with(prey))
  
  rho_E_2<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = prey,target = jonah_col, maxE = 7, showPlot = FALSE)
  E_out_2<-rho_E_2[which.max(rho_E_1$rho),"E"][1]
  out_2 <- CCM(dataFrame= df_temp, columns=prey, target=jonah_col, E = E_out_2, Tp=1,
               libSizes = paste(nrow(df_temp)-E_out_2, nrow(df_temp)-E_out_2, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE)  %>%
    mutate(prey=prey,direction= paste("prey","->","jonah"),
           E = E_out_2,stratum=stratum, trial_num = surr_trial) %>% 
    rename_with( ~ paste0("jonah:", prey, recycle0 = TRUE), starts_with("jonah")) %>% 
    rename_with( ~ paste0(prey, ":jonah",recycle0 = TRUE), starts_with(prey))
  

  bind_rows(out_1,out_2)
}) %>% mutate("jonah"=NA)  %>% addDirection()
```


```{r}
find_p_strat <- function(x,y) {
  ndist = null_dist(df = RESULTS_ccm_by_strat_surr %>% filter(stratum==x), s=y)
    val = RESULTS_ccm_by_strat_max %>% filter(stratum==x, name==y) %>% pull(value)
    prob = (sum(ndist >= val))/200
    return(prob)
}

strat_df <- data.frame(expand.grid(c(1:4),combos, stringsAsFactors = FALSE))
strat_df %>% rowwise() %>% mutate(p = find_p_strat(Var1, Var2))
```

### Weight
```{r}
#Create surrogate data
set.seed(7)
ccm_strat_wt_surr <- wtCCMdf_strat %>% group_by(Stratum) %>% mutate(new=surrogate(jonah, ns=200)) %>% select(-jonah) %>% mutate(new=data.frame(new)) %>% unnest(cols=new)

ccm_strat_wt_surr <- rename_with(ccm_strat_wt_surr,.cols=!any_of(c("date", "scallop", "rock", "Stratum")),~ paste0("jonah_", .x, recycle0 = TRUE) %>% str_remove("X"))
```

```{r}
RESULTS_ccm_wt_by_strat_surr <- pmap_dfr(params_strat_combos_surr,function(prey, stratum, surr_trial){
  df_temp <- ccm_strat_wt_surr %>% filter(Stratum==stratum)
  lib_vec <- paste(1, nrow(df_temp))
  jonah_col = paste("jonah",surr_trial, sep="_" )
  
  rho_E_1<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = jonah_col,target = prey, maxE = 7, showPlot = FALSE)
  E_out_1<-rho_E_1[which.max(rho_E_1$rho),"E"][1]
  out_1 <- CCM(dataFrame= df_temp, columns=jonah_col, target=prey, E = E_out_1, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_1, nrow(df_temp) - E_out_1, 
                                "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE) %>%
    mutate(prey=prey, direction= paste("jonah","->","prey"),
           E = E_out_1,stratum=stratum, trial_num = surr_trial) %>% 
    rename_with( ~ paste0("jonah:", prey, recycle0 = TRUE), starts_with("jonah")) %>% 
    rename_with( ~ paste0(prey, ":jonah",recycle0 = TRUE), starts_with(prey))
  
  rho_E_2<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = prey,target = jonah_col, maxE = 7, showPlot = FALSE)
  E_out_2<-rho_E_2[which.max(rho_E_1$rho),"E"][1]
  out_2 <- CCM(dataFrame= df_temp, columns=prey, target=jonah_col, E = E_out_2, Tp=1,
               libSizes = paste(nrow(df_temp)-E_out_2, nrow(df_temp)-E_out_2, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE)  %>%
    mutate(prey=prey,direction= paste("prey","->","jonah"),
           E = E_out_2,stratum=stratum, trial_num = surr_trial) %>% 
    rename_with( ~ paste0("jonah:", prey, recycle0 = TRUE), starts_with("jonah")) %>% 
    rename_with( ~ paste0(prey, ":jonah",recycle0 = TRUE), starts_with(prey))
  

  bind_rows(out_1,out_2)
}) %>% mutate("jonah"=NA)  %>% addDirection()
```

```{r}
find_p_strat_wt <- function(x,y) {
  ndist = null_dist(df = RESULTS_ccm_wt_by_strat_surr %>% filter(stratum==x), s=y)
    val = RESULTS_ccm_wt_by_strat_max %>% filter(stratum==x, name==y) %>% pull(value)
    prob = (sum(ndist >= val))/200
    return(prob)
}

strat_df %>% rowwise() %>% mutate(p = find_p_strat_wt(Var1, Var2))
```

## By area (done)

### Catch
```{r}
#Create the data frame that will store the combinations we want to test
params_areas_combos_surr <- map_dfr(seq_len(200), ~params_areas_ccm_combos %>% select(-"jonah"))
params_areas_combos_surr <- params_areas_combos_surr %>% mutate(surr_trial = rep(1:200, each=40))

#Create surrogate data
set.seed(7)
ccm_areas_surr <- catchCCMdf %>% group_by(area) %>% mutate(new=surrogate(jonah, ns=200)) %>% select(-c("jonah", "Region", "Stratum")) %>% mutate(new=data.frame(new)) %>% unnest(cols=new)

ccm_areas_surr <- rename_with(ccm_areas_surr,.cols=!any_of(c("date", "scallop", "rock", "area")),~ paste0("jonah_", .x, recycle0 = TRUE) %>% str_remove("X"))

```

```{r, eval=FALSE}
RESULTS_ccm_by_area_surr <- pmap_dfr(params_areas_combos_surr,function(prey, areaInput, surr_trial){
  df_temp <- ccm_areas_surr %>% filter(area==areaInput)
  lib_vec <- paste(1, nrow(df_temp))
  jonah_col = paste("jonah",surr_trial, sep="_" )
  
  rho_E_1<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = jonah_col,target = prey, maxE = 7, showPlot = FALSE)
  E_out_1<-rho_E_1[which.max(rho_E_1$rho),"E"][1]
  out_1 <- CCM(dataFrame= df_temp, columns=jonah_col, target=prey, E = E_out_1, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_1, nrow(df_temp) - E_out_1, 
                                "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE) %>%
    mutate(prey=prey, direction= paste("jonah","->","prey"),
           E = E_out_1,areaInput=areaInput, trial_num = surr_trial) %>% 
    rename_with( ~ paste0("jonah:", prey, recycle0 = TRUE), starts_with("jonah")) %>% 
    rename_with( ~ paste0(prey, ":jonah",recycle0 = TRUE), starts_with(prey))
  
  rho_E_2<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = prey,target = jonah_col, maxE = 7, showPlot = FALSE)
  E_out_2<-rho_E_2[which.max(rho_E_1$rho),"E"][1]
  out_2 <- CCM(dataFrame= df_temp, columns=prey, target=jonah_col, E = E_out_2, Tp=1,
               libSizes = paste(nrow(df_temp)-E_out_2, nrow(df_temp)-E_out_2, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE)  %>%
    mutate(prey=prey,direction= paste("prey","->","jonah"),
           E = E_out_2,areaInput=areaInput, trial_num = surr_trial) %>% 
    rename_with( ~ paste0("jonah:", prey, recycle0 = TRUE), starts_with("jonah")) %>% 
    rename_with( ~ paste0(prey, ":jonah",recycle0 = TRUE), starts_with(prey))
  

  bind_rows(out_1,out_2)
}) %>% mutate("jonah"=NA)  %>% addDirection()
```


```{r}
find_p_area <- function(x,y) {
  ndist = null_dist(df = RESULTS_ccm_by_area_surr %>% filter(areaInput==x), s=y)
    val = RESULTS_ccm_by_area_max %>% filter(areaInput==x, name==y) %>% pull(value)
    prob = (sum(ndist >= val))/200
    return(prob)
}


area_df <- data.frame(expand.grid(areaList,combos, stringsAsFactors = FALSE))
print(area_df %>% rowwise() %>% mutate(p = find_p_area(Var1, Var2)) %>% filter(p<0.1))
```

### Weight
```{r}
#Create surrogate data
set.seed(7)
ccm_wt_areas_surr <- wtCCMdf %>% group_by(area) %>% mutate(new=surrogate(jonah, ns=200)) %>% select(-c("jonah", "Region", "Stratum")) %>% mutate(new=data.frame(new)) %>% unnest(cols=new)

ccm_wt_areas_surr <- rename_with(ccm_wt_areas_surr,.cols=!any_of(c("date", "scallop", "rock", "area")),~ paste0("jonah_", .x, recycle0 = TRUE) %>% str_remove("X"))
```


```{r, eval=FALSE}
RESULTS_ccm_wt_by_area_surr <- pmap_dfr(params_areas_combos_surr,function(prey, areaInput, surr_trial){
  df_temp <- ccm_wt_areas_surr %>% filter(area==areaInput)
  lib_vec <- paste(1, nrow(df_temp))
  jonah_col = paste("jonah",surr_trial, sep="_" )
  
  rho_E_1<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = jonah_col,target = prey, maxE = 7, showPlot = FALSE)
  E_out_1<-rho_E_1[which.max(rho_E_1$rho),"E"][1]
  out_1 <- CCM(dataFrame= df_temp, columns=jonah_col, target=prey, E = E_out_1, Tp=1,
               libSizes = paste(nrow(df_temp) - E_out_1, nrow(df_temp) - E_out_1, 
                                "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE) %>%
    mutate(prey=prey, direction= paste("jonah","->","prey"),
           E = E_out_1,areaInput=areaInput, trial_num = surr_trial) %>% 
    rename_with( ~ paste0("jonah:", prey, recycle0 = TRUE), starts_with("jonah")) %>% 
    rename_with( ~ paste0(prey, ":jonah",recycle0 = TRUE), starts_with(prey))
  
  rho_E_2<- EmbedDimension(dataFrame = df_temp, lib = lib_vec, pred = lib_vec,
                           columns = prey,target = jonah_col, maxE = 7, showPlot = FALSE)
  E_out_2<-rho_E_2[which.max(rho_E_1$rho),"E"][1]
  out_2 <- CCM(dataFrame= df_temp, columns=prey, target=jonah_col, E = E_out_2, Tp=1,
               libSizes = paste(nrow(df_temp)-E_out_2, nrow(df_temp)-E_out_2, "1",sep=" "), sample=1, verbose=FALSE, showPlot = FALSE)  %>%
    mutate(prey=prey,direction= paste("prey","->","jonah"),
           E = E_out_2,areaInput=areaInput, trial_num = surr_trial) %>% 
    rename_with( ~ paste0("jonah:", prey, recycle0 = TRUE), starts_with("jonah")) %>% 
    rename_with( ~ paste0(prey, ":jonah",recycle0 = TRUE), starts_with(prey))
  

  bind_rows(out_1,out_2)
}) %>% mutate("jonah"=NA)  %>% addDirection()
```


```{r}
find_p_area_wt <- function(x,y) {
  ndist = null_dist(df = RESULTS_ccm_wt_by_area_surr %>% filter(areaInput==x), s=y)
    val = RESULTS_ccm_wt_by_area_max %>% filter(areaInput==x, name==y) %>% pull(value)
    prob = (sum(ndist >= val))/200
    return(prob)
}

print(area_df %>% rowwise() %>% mutate(p = find_p_area_wt(Var1, Var2)) %>% filter(p<0.1))
```

# Temperature

## By area