---
title: 'June 12'
date: '2023-06-12'
categories: ['Week 2', 'Data exploration', 'Reading']
description: 'My activities for the day.'
execute: 
  message: false
  warning: false
editor_options: 
  chunk_output_type: console
---

### Reading EDM papers
This morning I read many EDM papers and updated the Intro to EDM page with brief notes/summaries of many of them. I also did some reading on the different kinds of error (process vs. observational), stochastic processes, and white and red noise.

### Python Practice
Michael sent us an email today with additional material to review before Wednesday's Python programming workshop. I went through the information he sent and practiced working with Jupyter notebook in the SCC so I will be ready for the workshop.

## Data exploration

I started some preliminary data exploration/cleaning, although I have many questions for Dr. Deyle about what certain columns mean and about the overall structure of the data.

Load packages
```{r}
library(ggplot2)
library(dplyr)

```

Looking at 22564_UNION_FSCS_SVCAT (one out of five .csv files in the folder shared by Dr. Deyle)
```{r}
#read data
svcat <- read.csv("~/Downloads/lab_notebook/data/22564_UNION_FSCS_SVCAT.csv")

#view the first six rows of data
head(svcat)

#different overviews of the data
summary(svcat)
glimpse(svcat)
```

```{r}
#filter to only species of interest - reduces it from 94632 observations to 21425
scallops <- svcat %>% 
  filter(SCIENTIFIC_NAME=="Placopecten magellanicus (sea scallop)" | 
           SCIENTIFIC_NAME =="Placopecten magellanicus (sea scallop clapper)")

scallops <- scallops %>% 
  mutate(name = SCIENTIFIC_NAME,
         cruise6 = as.factor(CRUISE6),
         cruise = as.factor(CRUISE),
         sex = as.factor(CATCHSEX),
         stratum = as.factor(STRATUM),
         .keep = "unused")

summary(scallops)
```

```{r echo=FALSE}
#rm(svcat)
```

```{r}
scallops <- scallops %>% 
  mutate(name = recode(name, "Placopecten magellanicus (sea scallop)" = "scallop", "Placopecten magellanicus (sea scallop clapper)" = "clapper"))

summarise_all(scallops,n_distinct)

#All have sex = 0, means unknown sex
#SVSPP is species, listed as either 401 (scallop) or 400 (clammer), redundant info to species

#Remove unnecessary variables
scallops <- scallops %>% 
  select(-c("sex", "STATUS_CODE", "SVSPP"))

#ggplot(scallops, aes(x = cruise)) + geom_bar()
#ggplot(scallops, aes(x = cruise6)) + geom_bar()
#ggplot(scallops, aes(x = stratum)) + geom_bar()
  
```

