---
title: 'Week 2 in R'
date: '2023-06-15'
categories: ['Week 2', 'R']
description: 'My activities for the week'
execute: 
  message: false
  warning: false
editor_options: 
  chunk_output_type: console
---

Note: I continually refined my code, so some of the older or now-unnecessary code cells below are set to "eval = FALSE" so that the code will still be visible but will not be evaluated and disrupt the output of later cells.

## Monday, June 12

**All set to eval = FALSE**

Load packages

```{r, eval = FALSE}
library(ggplot2)
library(dplyr)

```

Looking at 22564_UNION_FSCS_SVCAT (one out of five .csv files in the folder shared by Dr. Deyle)

```{r, eval = FALSE}
#read data
svcat <- read.csv("~/Downloads/lab_notebook/data/22564_UNION_FSCS_SVCAT.csv")

#view the first six rows of data
head(svcat)

#different overviews of the data
summary(svcat)
glimpse(svcat)
```

```{r, eval = FALSE}
#filter to only species of interest - reduces it from 94632 observations to 21425
scallops <- svcat %>% 
  filter(SCIENTIFIC_NAME=="Placopecten magellanicus (sea scallop)" | 
           SCIENTIFIC_NAME =="Placopecten magellanicus (sea scallop clapper)")

scallops <- scallops %>% 
  mutate(name = SCIENTIFIC_NAME,
         cruise6 = as.factor(CRUISE6),
         cruise = as.factor(CRUISE),
         sex = as.factor(CATCHSEX),
         stratum = as.factor(STRATUM),
         .keep = "unused")

summary(scallops)
```

```{r, eval = FALSE}
scallops <- scallops %>% 
  mutate(name = recode(name, "Placopecten magellanicus (sea scallop)" = "scallop", "Placopecten magellanicus (sea scallop clapper)" = "clapper"))

summarise_all(scallops,n_distinct)

#All have sex = 0, which means "Unsexed, unknown, or sex not observed"
#SVSPP is species, listed as either 401 (scallop) or 400 (clapper), redundant info to species

#Remove unnecessary variables
scallops <- scallops %>% 
  select(-c("sex", "STATUS_CODE", "SVSPP"))

#ggplot(scallops, aes(x = cruise)) + geom_bar()
#ggplot(scallops, aes(x = cruise6)) + geom_bar()
#ggplot(scallops, aes(x = stratum)) + geom_bar()
  
```

# Tuesday and Wednesday (June 13 and 14)

Load packages

```{r}
library(tidyverse) #rather than importing separately

```

Set-up (from June 12)

```{r}
#read data
svcat <- read.csv("~/Downloads/lab_notebook/data/Old/22564_UNION_FSCS_SVCAT.csv")

#filter to only species of interest - reduces it from 94632 observations to 21425
scallops <- svcat %>% 
  filter(SCIENTIFIC_NAME=="Placopecten magellanicus (sea scallop)" | 
           SCIENTIFIC_NAME =="Placopecten magellanicus (sea scallop clapper)")

#turn categorical variables into factors
scallops <- scallops %>% 
  mutate(name = SCIENTIFIC_NAME,
         CRUISE6 = as.factor(CRUISE6),
         CRUISE = as.factor(CRUISE),
         CATCHSEX = as.factor(CATCHSEX),
         STRATUM = as.factor(STRATUM),
         .keep = "unused")

#shorten for clarity
scallops <- scallops %>% 
  mutate(name = recode(name, "Placopecten magellanicus (sea scallop)" = "scallop", 
                       "Placopecten magellanicus (sea scallop clapper)" = "clapper"))

#Remove unnecessary/redundant variables: all entries are listed as unknown sex (sex=0), 
# all have the same status code (10), and SVSPP corresponds with name (401 = scallop, 
# 400 = clappers)
scallops <- scallops %>% 
  select(-c("CATCHSEX", "STATUS_CODE", "SVSPP"))
```

I worked on turning the data into an actual time series:

```{r}
#read the file that has the season and year associated with each cruise
cruises <- read.csv("~/Downloads/lab_notebook/data/Old/22564_SVDBS_CRUISES.csv")

#select the relevant variables
cruises <- cruises %>% 
  select(c("CRUISE6","SEASON","YEAR")) %>% 
  mutate(CRUISE6 = as.factor(CRUISE6), .keep="unused")

#combine with the existing dataframe
scallops <- left_join(scallops, cruises, by="CRUISE6")

as_tibble(scallops)

```

# Thursday (June 15)

Load packages

```{r, eval=FALSE}
library(tidyverse)
library(lubridate)

```

Dr. Deyle sent me some example code for wrangling lobster data from the NEFSC trawl surveys. I adapted the code for use with the scallop dredge surveys:

```{r}
#read the file that has the metadata associated with each stratum and station
df_strata <- read.csv("data/Old/SVDBS_SVMSTRATA.csv")
df_stations <- read.csv("data/Old/22564_UNION_FSCS_SVSTA.csv")

#standardizing the date formats
df_stations_wrangled <- df_stations %>%
  mutate(DATE = parse_date_time(BEGIN_EST_TOWDATE,orders="mdYHMS",truncated = 3)) %>%
  select(STATION,STRATUM,DATE) %>% #select relevant columns
  mutate(DATE=year(DATE)) %>%
  distinct() #remove duplicate rows

```

Then, I wanted to get a general idea of how many times each station and stratum were in the data (i.e., how many years that station or stratum was sampled).

```{r}
#shows how many occurrences of each station are in the data (i.e. how many years that station was sampled)
data.frame(table(df_stations_wrangled$STATION))
data.frame(table(df_stations_wrangled$STRATUM))
```

This revealed that 20 rows had no station recorded. Additionally, many rows had text of some sort rather than a station or stratum number recorded in the corresponding column (e.g., "took 20-30 meters" and "STARFISH NO STARS").


I then checked to see how many entries were present for each year:
```{r}
data.frame(table(df_stations_wrangled$DATE))
```

Importantly, this revealed that the dates for stations only go up to 2006, plus a few observations in 2015 and 2021. 
At first, I thought that this meant there was only scallop data until 2006, and that this dataset didn't include the SMAST and WHOI drop camera data that has been collected since around 2003 as part of the NEFSC scallop survey. The example code for wrangling lobster data relies on the fact that each lobster observation is associated with a station, so I assumed (incorrectly) that the scallop data would be the same.

Unfortunately, this prompted me to spend much of the afternoon searching for the HabCam data, until I eventually realized that only having stations until 2006 did not mean there was only scallop data until 2006. Checking the entries per year for the scallop data rather than the stations revealed that there was in fact scallop data from 1977 through 2021.

```{r}
data.frame(table(scallops$YEAR))
```

However, searching for the HabCam data was not a complete waste of time. In my search, I had re-pulled the NOAA Scallop Survey data and found that this newer version had data through 2022, whereas the version that Dr. Deyle shared with me only extended to 2021. See the "Week 2 in Words" page for more about what I found in my search for the post-2006 scallop data.

```{r}
#re-run setup for new data
svcatNew <- read.csv("~/Downloads/lab_notebook/data/new22564_UNION_FSCS_SVCAT.csv")

scallopsNew <- svcatNew %>% 
  filter(SCIENTIFIC_NAME=="Placopecten magellanicus (sea scallop)" | 
           SCIENTIFIC_NAME =="Placopecten magellanicus (sea scallop clapper)")

#turn categorical variables into factors
scallopsNew <- scallopsNew %>% 
  mutate(name = SCIENTIFIC_NAME,
         CRUISE6 = as.factor(CRUISE6),
         CRUISE = as.factor(CRUISE),
         CATCHSEX = as.factor(CATCHSEX),
       #  STRATUM = as.factor(STRATUM), <-- this became unnecessary later
         .keep = "unused")

#shorten for clarity
scallopsNew <- scallopsNew %>% 
  mutate(name = recode(name, "Placopecten magellanicus (sea scallop)" = "scallop", 
                       "Placopecten magellanicus (sea scallop clapper)" = "clapper")) %>% 
 select(-c("CATCHSEX", "STATUS_CODE", "SVSPP"))

#read the file that has the season and year associated with each cruise
cruisesNew <- read.csv("~/Downloads/lab_notebook/data/new22564_SVDBS_CRUISES.csv")

#select the relevant variables
cruisesNew <- cruisesNew %>% 
  select(c("CRUISE6","SEASON","YEAR")) %>% 
  mutate(CRUISE6 = as.factor(CRUISE6), .keep="unused")

#combine with the existing dataframe
scallopsNew <- left_join(scallopsNew, cruisesNew, by="CRUISE6")

as_tibble(scallopsNew)

data.frame(table(scallopsNew$YEAR))
  
```

I then started to pull data where there was all 45 years of data collection for a single stratum.

```{r, warning=FALSE}

#this gets rid of some weird strata in 1984 with letters, which don't seem to 
# correspond to any of the NOAA (shellfish, dredge, or trawl survey) strata
scallopsTest <-  scallopsNew %>%
  mutate(STRATUM = as.integer(STRATUM)) %>% 
  filter(!is.na(STRATUM))

#This will show you which ones were removed
# scallopsTest <-  scallopsNew %>%
#   mutate(strat = as.integer(STRATUM)) %>% 
#   filter(is.na(strat))

df_strataNew <- read.csv("data/newSVDBS_SVMSTRATA.csv")

#shows how many occurrences of each stratum are in the data (i.e. how many years that stratum was sampled)
data.frame(table(scallopsNew$STRATUM))

#the key matching strata to location (latitude and longitude) starts from 1010
scallopsNew <-  scallopsNew %>%
  filter(STRATUM > 1000)

```

