

Now, we're going to take a different approach. Instead of utilizing the
different scales of spatial replication by running CCM on them
individually to identify similarities and differences, we're going to do
multispatial CCM, where we treat each spatial area as a replicate and
concatenate the areas to create a longer time series, discarding rows
where lags cross time series. This will give us 3 more versions of CCM
to investigate overall system dynamics:

5.  Multispatial CCM (aggregate) with areas as replicates = run CCM on 1
    \>500-point time series
6.  Multispatial CCM (aggregate) with regions (averaged across strata)
    as replicates = run CCM on 1 \~150-point time series
7.  Multispatial CCM (aggregate) with strata (averaged across regions)
    as replicates = run CCM on 1 \~115-point time series



Let's jump into the multispatial CCM.

```{r}
#There are the columns that we will want to keep from the CCM output
v_keep_col <- c("E","Tp","num_pred", "rho", "mae","rmse","perc","p_val","rho_linear", "mae_linear",
                "rmse_linear","perc_linear","p_val_linear")
```

First, we'll need a basic delay function:

```{r}
delay <- function(x,n){
  if(n>=0) 
    lead(x,n) 
  else 
    lag(x,abs(n))}
```

Now, we need to make the embedding block for the CCM. The rEDM package
has a function ("MakeBlock") that does this, but it cannot handle
embedding concatenated replicates, since we can't lag the beginning of
one replicate's time series based on the end of another replicate's time
series. Instead, we create a function that lags after grouping by an ID
column, then binding the individually-lagged replicates. Incomplete rows
will then conveniently have NAs in them, making them easy to identify
and remove.

```{r}
# Function 2B: make_xmap_block_ID ---------------------------------------------

make_xmap_block_ID <- function(df,predictor,target,ID_col,E_max,cause_lag){
  
  v_delays <- 0:-(E_max-1)
  
  df_ccm <- df %>%
    select({{ID_col}},{{target}}) %>%
    group_by({{ID_col}}) %>%
    transmute(target=delay({{target}},cause_lag))
  
  df_lags <- map_dfc(v_delays,function(d_i){
    df %>%
      group_by({{ID_col}}) %>%
      transmute("pred_t{d_i}" := delay({{predictor}},d_i)) %>%
      ungroup({{ID_col}}) %>% select(-{{ID_col}})
  })
  
  df_out <- bind_cols(df_ccm,df_lags) %>% 
    ungroup() %>%
    mutate(index=row_number()) %>%
    select(index,everything())
  
  return(df_out)
  
}
```

Now we will define a helper function to make the multispatial CCM more
efficient. For consistency with the normal CCM we performed earlier, we
will use E=2.

```{r}

do_xmap_ID <- function(df,predictor,target,ID_col,E,tp=1){

  df_1 <- make_xmap_block_ID(df,!!sym(predictor),!!sym(target),!!sym(ID_col),
                          E_max=E,cause_lag=tp) %>% ungroup() %>% filter(complete.cases(.))
  
  lib_1 <- paste(1,nrow(df_1))
 
  columns_star <- names(df_1)[4:(E+3)]

  out <- Simplex(dataFrame=df_1,
                 lib=lib_1,pred=lib_1,Tp=0, # The target has already been "manually" lagged
                 target="target",
                 columns=columns_star,
                 embedded=TRUE,
                 parameterList = TRUE,
                 E=E)

  fit_linear <- lm(as.formula(paste0("target ~ ",columns_star[1])),data=df_1)
  out_linear_predictions <- predict(fit_linear)

  params <- out$parameters
  out <- out$predictions %>% filter(complete.cases(.))

  stats <- compute_stats(out$Observations,out$Predictions)

  stats_linear <- compute_stats(out$Observations,out_linear_predictions)
  names(stats_linear) <- paste0(names(stats_linear),"_linear")

    return(bind_cols(Filter(function(x) length(x)==1,params),stats,stats_linear))
}
```

Let's test our functions with some toy data:

```{r}
data("sardine_anchovy_sst")
set.seed(7)

test_block1 <- sardine_anchovy_sst %>% select(year, anchovy, np_sst) %>% slice(1:35) %>% mutate(ID = 1)
test_block2 <- test_block1 %>% mutate(across(anchovy:np_sst, ~jitter(.x*2, factor=1))) %>% mutate(ID = 2)
test_1 <- rbind(test_block1, test_block2)

test_block <- make_xmap_block_ID(test_1,target=np_sst,predictor = anchovy,ID_col = ID,E_max=3,cause_lag=1) 
test_block %>% group_by(ID) %>% slice_head(n=5) #works as expected

#Test with only one replicate
do_xmap_ID(df=test_block1,predictor = "anchovy", target="np_sst",ID_col = "ID", E=3) %>% select(all_of(v_keep_col))

CCM(dataFrame = sardine_anchovy_sst %>% slice(1:35), E = 3, Tp = 1, columns = "anchovy", target = "np_sst", libSizes = "32 32 1", sample = 100) #Traditional CCM gives essentially the same answer— it is unable to recover the true causal interactions from the short time series

#Test with the two replicates, the second being a slightly altered version of the first
do_xmap_ID(df=test_1,predictor = "anchovy", target="np_sst",ID_col = "ID", E=3) %>% select(all_of(v_keep_col))

do_xmap_ID(df=test_1,predictor = "np_sst", target="anchovy",ID_col = "ID", E=3) %>% select(all_of(v_keep_col))

CCM(dataFrame = sardine_anchovy_sst, E = 3, Tp = 1, columns = "anchovy", target = "np_sst", libSizes = "70 70 1", sample = 100, seed=7) #compare with normal CCM on the 78-year data set with a library size of 70, the same length as our test_1 concatenated time series


```

Our functions appear to work as we want them to. The embedded block is
constructed correctly, with the NAs where appropriate. Originally when
we select only 35 observations of the data, the predictive skill is
negative for our do_xmap_ID function, which is essentially doing exactly
the same Simplex call as normal CCM and equally struggling with the
short time series.

However, when we concatenate the replicates, the predictive skill vastly
improves and we can identify the true causal relationship between sea
surface temperature and anchovy abundance, with a positive rho for
anchovy->SST and a negative rho for SST->anchovy. Since the
multispatial model can use the past states of replicate 1 from time
point t when it is predicting the same time point t for replicate 2 with
leave-one-out cross-validation (and vice versa), adding a slightly
different version of the 35 time point data as a replicate brought
cross-map skill to an even higher rho than normal CCM for anchovy->SST
with a library size of 70.

If we change the second replicate to be less similar to the first, the predictive skill should decrease
```{r}
# Multiply the original by 3 instead of 2 and increase factor from 1 to 3
test_block3 <- test_block1 %>% mutate(across(anchovy:np_sst, ~jitter(.x*3, factor=3))) %>% mutate(ID = 2)
test_2 <- rbind(test_block1, test_block3)

#Test with the two replicates, the second being a slightly altered version of the first
do_xmap_ID(df=test_2,predictor = "anchovy", target="np_sst",ID_col = "ID", E=3) %>% select(all_of(v_keep_col))

```

Rho decreased from 0.38 to 0.13, reflecting the fact that including overly heterogeneous replicates may not improve model performance.

We'll start with multispatial CCM using areas as replicates. For simplicity and consistency with our original aggregate CCM, we will use an embedding dimension of 2.

```{r}
# Areas as replicates - catch
RESULTS_ccm_combos_areas <- pmap_dfr(params_ccm_combos,function(jonah,prey){
  
  out_1 <- do_xmap_ID(catchCCMdf,predictor=jonah,target=prey,ID_col = "area",E = 2, tp=1) %>% 
    mutate(direction= paste("jonah","->",prey))
  
  out_2 <- do_xmap_ID(catchCCMdf,predictor=prey,target=jonah,ID_col = "area",E = 2, tp=1) %>% 
    mutate(direction= paste(prey,"->","jonah"))
  
  bind_rows(out_1,out_2) %>% select(direction, all_of(v_keep_col))
  
}) %>% mutate(type="catch", replicate = "area")

# Areas as replicates - weight
RESULTS_ccm_wt_combos_areas <- pmap_dfr(params_ccm_combos,function(jonah,prey){
  
  out_1 <- do_xmap_ID(wtCCMdf,predictor=jonah,target=prey,ID_col = "area",E = 2, tp=1) %>% 
    mutate(direction= paste("jonah","->",prey))
  
  out_2 <- do_xmap_ID(wtCCMdf,predictor=prey,target=jonah,ID_col = "area",E = 2, tp=1) %>% 
    mutate(direction= paste(prey,"->","jonah"))
  
  bind_rows(out_1,out_2) %>% select(direction, all_of(v_keep_col))  %>% mutate(type="wt", replicate = "area")
  
})

```

Now we move on to using regions as replicates
```{r}
regionsDf<- catchCCMdf %>% ungroup() %>% 
  group_by(Region, date) %>% 
  summarise(across(scallop:jonah, ~ mean(.x, na.rm = TRUE)))

regionsDf_wt<- wtCCMdf %>% ungroup() %>% 
  group_by(Region, date) %>% 
  summarise(across(scallop:jonah, ~ mean(.x, na.rm = TRUE)))

# Regions as replicates
RESULTS_ccm_combos_regions <- pmap_dfr(params_ccm_combos,function(jonah,prey){
  
  out_1 <- do_xmap_ID(regionsDf,predictor=jonah,target=prey,ID_col="Region",E=2, tp=1)  %>%
    mutate(direction= paste("jonah","->",prey))
  
  out_2 <- do_xmap_ID(regionsDf,predictor=prey,target=jonah,ID_col="Region",E=2, tp=1) %>% 
    mutate(direction= paste(prey,"->","jonah"))
  
  bind_rows(out_1,out_2) %>% select(direction, all_of(v_keep_col))  %>% mutate(type="catch", replicate = "region")
  
})

# Regions as replicates
RESULTS_ccm_wt_combos_regions <- pmap_dfr(params_ccm_combos,function(jonah,prey){
  
  out_1 <- do_xmap_ID(regionsDf_wt,predictor=jonah,target=prey,ID_col="Region",E=2, tp=1)  %>%
    mutate(direction= paste("jonah","->",prey))
  
  out_2 <- do_xmap_ID(regionsDf_wt,predictor=prey,target=jonah,ID_col="Region",E=2, tp=1) %>% 
    mutate(direction= paste(prey,"->","jonah"))
  
  bind_rows(out_1,out_2) %>% select(direction, all_of(v_keep_col))  %>% mutate(type="wt", replicate = "region")
})


```

And strata:
```{r}
strataDf<- catchCCMdf %>% ungroup() %>% 
  group_by(Stratum, date) %>% 
  summarise(across(scallop:jonah, ~ mean(.x, na.rm = TRUE)))

# Strata as replicates
RESULTS_ccm_combos_strata <- pmap_dfr(params_ccm_combos,function(jonah,prey){
  
  out_1 <- do_xmap_ID(strataDf,predictor=jonah,target=prey,ID_col="Stratum",E=2, tp=1)  %>%
    mutate(direction= paste("jonah","->",prey))
  
  out_2 <- do_xmap_ID(strataDf,predictor=prey,target=jonah,ID_col="Stratum",E=2, tp=1) %>% 
    mutate(direction= paste(prey,"->","jonah"))
  
  bind_rows(out_1,out_2) %>% select(direction, all_of(v_keep_col))  %>% mutate(type="catch", replicate = "stratum")
  })
  
#WEIGHT
strataDf_wt<- wtCCMdf %>% ungroup() %>% 
  group_by(Stratum, date) %>% 
  summarise(across(scallop:jonah, ~ mean(.x, na.rm = TRUE)))

# Strata as replicates - weight
RESULTS_ccm_wt_combos_strata <- pmap_dfr(params_ccm_combos,function(jonah,prey){
  
  out_1 <- do_xmap_ID(strataDf_wt,predictor=jonah,target=prey,ID_col="Stratum",E=2, tp=1)  %>%
    mutate(direction= paste("jonah","->",prey))
  
  out_2 <- do_xmap_ID(strataDf_wt,predictor=prey,target=jonah,ID_col="Stratum",E=2, tp=1) %>% 
    mutate(direction= paste(prey,"->","jonah"))
  
  bind_rows(out_1,out_2) %>% select(direction, all_of(v_keep_col)) %>% mutate(type="wt", replicate = "stratum")
  
})
```

Let's compare
```{r}
rbind(RESULTS_ccm_combos_areas,
      RESULTS_ccm_combos_regions,
      RESULTS_ccm_combos_strata,
      RESULTS_ccm_wt_combos_areas,
      RESULTS_ccm_wt_combos_regions,
      RESULTS_ccm_wt_combos_strata) %>% select(c("type", "replicate", "direction", "num_pred", "rho", "rho_linear", "p_val"))
```

These are all pretty bad, only rarely doing better than the linear model and never exceeding a rho of 0.24. The clear spatial heterogeneity we saw earlier may mean that the manifolds represented by the different replicates are too dissimilar to yield accurate predictions when combined.