---
title: 'Methods'
date: '2023-07-13'
execute: 
  message: false
  warning: false
editor_options: 
  chunk_output_type: console
#bibliography: references.bib
---

Ruby Krasnow
Last modified: July 13, 2023

Load packages
```{r}
library(tidyverse)
library(lubridate) #date formatting
library(patchwork) #combining plots
library(tseries) #for KPSS test for stationarity
library(rEDM) #EDM

#spatial packages
library(sf)
library(sfheaders)
library(spdep)
```

```{r, include=FALSE}
sys.source("~/Downloads/lab_notebook/Maine/helpFunctionsMaine.R", envir = knitr::knit_global())
```

# Organize data
Analysis was performed on the catch weight (kg) of each species, which was recorded for each tow and standardized by DMR to account for tows that were longer or shorter than the 1 NM standard. 
```{r}

df_tows<-read.csv("~/Downloads/lab_notebook/data/Maine_inshore_trawl/MEtows.csv") #tow data
df_s_cat<- read.csv("~/Downloads/lab_notebook/data/Maine_inshore_trawl/MEscallopCatch.csv") #scallop catch
df_r_cat<- read.csv("~/Downloads/lab_notebook/data/Maine_inshore_trawl/MErockCatch.csv") #rock crab catch
df_j_cat<- read.csv("~/Downloads/lab_notebook/data/Maine_inshore_trawl/MEjonahCatch.csv") #jonah crab catch

```

## Log-transform, summarize, and assess stationarity
All data was log transformed to reduce nonstationarity, which was assessed for each study area using the Kwiatkowski-Phillips-Schmidt-Shin (KPSS) test (Kwiatkowski et al., 1992).
```{r}
# cleanCatch function includes log transformation; see helpFunctionsMaine script for details

s_cat_clean <- cleanCatch(df_s_cat) %>% 
  mutate(Common_Name = "Scallop")

r_cat_clean <- cleanCatch(df_r_cat) %>% 
  mutate(Common_Name = "Rock")

j_cat_clean <- cleanCatch(df_j_cat) %>% 
  mutate(Common_Name = "Jonah")

#Reorder columns
colOrder<-c("area", "Survey", "Tow_Number", "Region", "Stratum", "logCatch", "logWt", "Expanded_Catch", 
  "Expanded_Weight_kg", "Date", "Common_Name", "Number_Caught", "Weight_kg",
  "Start_Latitude", "Start_Longitude","Season",
  "Year","Grid", "Start_Depth_fathoms", "End_Depth_fathoms",
  "Bottom_WaterTemp_DegC", "Bottom_Salinity")

j_cat_clean <- j_cat_clean %>% select(all_of(colOrder))
r_cat_clean <- r_cat_clean %>% select(all_of(colOrder))
s_cat_clean <- s_cat_clean %>% select(all_of(colOrder))

#computes averages for each study area (area = region-stratum combination)
j_cat_sum <- summaryCatch(j_cat_clean)
r_cat_sum <- summaryCatch(r_cat_clean)
s_cat_sum <- summaryCatch(s_cat_clean)

catch <- s_cat_sum %>% left_join(j_cat_sum, by=c("area", "Season", "Region", "Stratum", "Year"), suffix = c("_s", "_j"))
catch <- catch %>% left_join(r_cat_sum, by=c("area", "Season", "Region", "Stratum", "Year")) %>% 
  mutate(avgCatch_r = avgCatch,avgWt_r = avgWt,avgLogCatch_r = avgLogCatch,avgLogWt_r = avgLogWt, .keep="unused")

catchTidy <- pivot_longer(catch, 
        cols = 6:ncol(catch)) %>% 
  mutate(Type = case_when(
    startsWith(name, "avgCatch_") ~"catch",
    startsWith(name,"avgWt_") ~"wt",
    startsWith(name,"avgLogWt") ~"logWt",
    startsWith(name,"avgLogCatch") ~"logCatch")) %>% 
  mutate(Species = case_when(
    endsWith(name, "s") ~"scallop",
    endsWith(name, "r") ~"rock",
    endsWith(name, "j") ~"jonah"))

catchTidy <- catchTidy %>% mutate(area = as.factor(area), Species = as.factor(Species),Season = as.factor(Season),Region = as.factor(Region), Stratum = as.factor(Stratum)) %>% 
          select(-name)

logCatchFall <- catchTidy %>% filter(Type == "logCatch") %>% filter(Season == "Fall")
logWtFall <- catchTidy %>% filter(Type == "logWt") %>% filter(Season == "Fall")

#Note that the highest p-value the kpss.test function will show is 0.1, actual p-values may be higher
findSpeciesKPSS(logCatchFall %>% filter(Species=="scallop"), season="Fall", type="logCatch")
findSpeciesKPSS(logCatchFall %>% filter(Species=="rock"), season="Fall", type="logCatch")
findSpeciesKPSS(logCatchFall %>% filter(Species=="jonah"), season="Fall", type="logCatch")

findSpeciesKPSS(logWtFall %>% filter(Species=="scallop"), season="Fall", type="logWt")
findSpeciesKPSS(logWtFall %>% filter(Species=="rock"), season="Fall", type="logWt")
findSpeciesKPSS(logWtFall %>% filter(Species=="jonah"), season="Fall", type="logWt")


```

# Question 1: Do the intensity and pattern of spatial autocorrelation indicate spatial bootstrapping could be an appropriate method?

First, import the location data for the survey region (surveyGrid) and transform the catch dataframes into sf objects for spatial analysis, using the same CRS as the survey grid.
```{r}
surveyGrid <-st_read("~/Downloads/lab_notebook/Maine/MaineDMR_-_Inshore_Trawl_Survey_Grid/") 
#CRS: WGS 84/EPSG 4326

surveyGrid <- surveyGrid %>% 
  mutate(Region = region_id,
         Stratum = depth_stra,
         GridID = grid_id, .keep="unused", .before=last_surve)

surveyGrid$region_stratum <- paste(surveyGrid$Region, surveyGrid$Stratum)

ggplot(surveyGrid)+geom_sf()

s_cat_sf <- st_as_sf(s_cat_clean, coords = c("Start_Longitude", "Start_Latitude"), crs=4326)
r_cat_sf <- st_as_sf(r_cat_clean, coords = c("Start_Longitude", "Start_Latitude"), crs=4326)
j_cat_sf <- st_as_sf(j_cat_clean, coords = c("Start_Longitude", "Start_Latitude"), crs=4326)

# Map of all points over grid
ggplot() + geom_sf(data = surveyGrid) + geom_sf(data = s_cat_sf)

```

Merge the individual survey grids into one region
```{r}
mergedGrid <- st_union(surveyGrid, by_feature = FALSE) %>% st_sf()
mergedGridBuffer <- st_buffer(mergedGrid, 9000)
ggplot(mergedGridBuffer)+geom_sf()
```

First, Voronoi polygons were constructed around each data point.
```{r, message=FALSE}
voronoiScal <- s_cat_sf %>%
  st_geometry() %>% # to get sfc from sf
  st_union() %>% # to get a sfc of MULTIPOINT type
  st_voronoi() %>% #
  st_collection_extract(type = "POLYGON") %>% # a list of polygons
  st_sf() %>% # from list to sf object
  st_intersection(mergedGridBuffer) %>%
  st_join(s_cat_sf)  # put names back

plot(voronoiScal$geometry)
```


Calculate Global Moran's I
```{r}
#Create a list of neighbors for each polygon
nb <- poly2nb(voronoiScal, queen = TRUE) # queen shares point or border
nbw <- nb2listw(nb, style = "W") #creates a weighted list (e.g., if a polygon has 4 neighbors, each neighbor receives a weight of 0.25)

# Global Moran's I - used more robust Monte Carlo method rather than the analytical method
gmoranScal <- moran.mc(voronoiScal$logCatch, nbw, nsim=999, alternative="greater")
gmoranScal

gmoranScalWt <- moran.mc(voronoiScal$logWt, nbw, nsim=999, alternative="greater")
gmoranScalWt

```


Calculate and display Local Moran's I

```{r}
lmoran <- localmoran(voronoiScal$logCatch, nbw, alternative = "greater")
head(lmoran)

lmoranWt <- localmoran(voronoiScal$logWt, nbw, alternative = "greater")
head(lmoranWt)

voronoiScal$lmI <- lmoran[, "Ii"] # local Moran's I - catch
voronoiScal$lmIwt <- lmoranWt[, "Ii"] # local Moran's I - weight
voronoiScal$lmZ <- lmoran[, "Z.Ii"] # z-scores - catch
voronoiScal$lmZwt <- lmoranWt[, "Z.Ii"] # z-scores - weight

# p-values corresponding to alternative greater
voronoiScal$lmp <- lmoran[, "Pr(z > E(Ii))"] #catch
voronoiScal$lmpWt <- lmoranWt[, "Pr(z > E(Ii))"] #weight

catchPlot <- ggplot(voronoiScal)+geom_sf(aes(fill=logCatch))+scale_fill_viridis_c(option = "D")+theme_bw()
moranPlotZ <- ggplot(voronoiScal)+geom_sf(aes(fill=lmZ))+scale_fill_viridis_c(option = "G", name="Local Moran's I\nZ-score (catch)")+theme_bw()

wtPlot <- ggplot(voronoiScal)+geom_sf(aes(fill=logWt))+scale_fill_viridis_c(option = "D")+theme_bw()
moranPlotWtZ <- ggplot(voronoiScal)+geom_sf(aes(fill=lmZwt))+scale_fill_viridis_c(option = "G", name="Local Moran's I\nZ-score (weight)")+theme_bw()

catchPlot + moranPlotZ + wtPlot + moranPlotWtZ + plot_annotation(tag_levels = 'A') + plot_layout(tag_level = 'new') & 
  theme(plot.tag = element_text(size = 12, vjust = 6))
```

# Question 2: Does spatial replication using neighboring points preserve the dimensionality and nonlinearity of a system?

```{r}
# calculate the average number of tows per area over the whole time series
# there are a few years where there is one more tow for scallops than for the crab species
numTows<- (s_cat_clean %>% group_by(area) %>% summarise(num = n_distinct(row_number())))
numTowsCrabs<- (r_cat_clean %>% group_by(area) %>% summarise(num = n_distinct(row_number())))
```

